---
title: "Homework 2"
output: github_document
---
# Problem 1

* Mr. Trash Wheel Sheet
```{r}
#Load tidyverse library
library(tidyverse)

# Import in the Mr. Trash Wheel sheet from mrtrashwheel dataset
# First row is skipped as it contains company logo, blank cells are
# dropped, and the names are cleaned up
mrtrashwheel_data = 
  readxl::read_excel(
    "./data_hw2/trash_wheel_update.xlsx", 
    sheet = 1, 
    skip = 1,
    range = "A2:N338",
    col_names = TRUE,
    col_types = NULL) %>%
janitor::clean_names() %>%
#Drop any blank cells
drop_na() %>%
#Select everything but homes_powered
select(-homes_powered) %>%
#Change sports_balls to an integer format
mutate(sport_balls = as.integer(sports_balls)) %>%
#Drop non-integer sports ball variable
select(-sports_balls)

mrtrashwheel_data

```

* Precipitation Sheets
```{r}
#Import precipitation 2017 data from the 3rd sheet, mutate year to 2017, 
# clean up the names, and drop any rows with blank cells
precipitation_2017 = 
  readxl::read_excel(
    "./data_hw2/trash_wheel_update.xlsx", 
    sheet = 6,
    range = "A2:B14",
    col_names = TRUE,
    col_types = NULL) %>%
  mutate(year = "2017") %>%
  janitor::clean_names() %>%
  mutate(month = month.name[as.numeric(month)]) %>%
  drop_na()

precipitation_2017

#Import precipitation data from 2018 from the 4th sheet, 
# mutate year to 2018, clean up the names, and drop any blank cells
precipitation_2018 = 
  readxl::read_excel(
    "./data_hw2/trash_wheel_update.xlsx", 
    sheet = 5,
    range = "A2:B14",
    col_names = TRUE,
    col_types = NULL) %>%
  mutate(year = "2018") %>%
  janitor::clean_names() %>%
  mutate(month = month.name[as.numeric(month)]) %>%
  drop_na()

precipitation_2018

#Combine the two datasets into one precipitation_data set, bind rows, clean the names, mutate the month to month name
precipitation_data = 
  bind_rows(precipitation_2017, precipitation_2018) %>%
  janitor::clean_names()

precipitation_data
```
The data in the mrtrashwheel_data set has `r nrow(mrtrashwheel_data)` number of observations. The data describe the contents, weight, and volume of dumpsters during 2015 from April to December. The mean weight of the dumpsters is `r mean(pull(mrtrashwheel_data, weight_tons))` and the median number of sports balls in a dumpster in 2017 is `r median(pull(filter(mrtrashwheel_data, year == 2017), sport_balls))`. The data in precipitation_data has `r nrow(precipitation_data)` number of observations. The data show the amount of precipitation per month in 2017 and 2018. For 2018 the amount of precipitation is `r sum(pull(filter(precipitation_data, year == 2018), total))` inches and for 2017 the amount of precipitation is `r sum(pull(filter(precipitation_data, year == 2017), total))` inches.

# Problem 2
*Clean data in pols-month.csv
```{r, echo = FALSE}
#Import in pols_month data
pols_month_data =
  read_csv(file = "./data_hw2/fivethirtyeight_datasets/pols-month.csv") %>%
  #Clean names
  janitor::clean_names() %>%
  #Separate the date into year, month, and day
  separate(mon, c("year", "month", "day"), convert = TRUE) %>%
   mutate(month = month.name[as.numeric(month)]) %>%
  #Pivot longer to make prez_gop and prez_dem one column
  pivot_longer(starts_with("prez_"),
               values_to = "binary") %>%
  #Filter for the ones in the binary column
  filter(binary == 1) %>%
  #Arrange by year and month
  arrange(year, month) %>%
  #Get rid of binary and day columns, change the column we pivoted to president
  select(-day,-binary, president = name) %>%
  #Transform president so that we drop "prez_" from the beginning of each value
  transform(president = str_replace(president, "prez_", ""))%>%
  #Make sure each row is distinct (remove duplicates)
  distinct()

pols_month_data

```

*Clean data in snp.csv
```{r}
#Import SNP data
snp_data =
  read_csv(file = "./data_hw2/fivethirtyeight_datasets/snp.csv") %>%
#Clean names
  janitor::clean_names() %>%
#Separate out date as month, day, and year
  separate(date, c("month","day","year"), convert = TRUE) %>%
#Change month to month name
  mutate(month = month.name[as.numeric(month)]) %>%
#Arrange the data with year and month as leading columns
  arrange(year, month) %>%
#Select everything, but the day column
  select(year, month, close, -day) 

snp_data
```


* Clean unemployment data
```{r}
#Import unemployment dataset
unemployment_data = 
  read_csv(file = "./data_hw2/fivethirtyeight_datasets/unemployment.csv") %>%
#Pivot the data so that Jan through Dec rows are columns with month as variable
   pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "rate") %>%
#Mutate the month to be month name
   mutate(month = month.name[match(month,month.abb)]) %>%
#Clean names
  janitor::clean_names() 
 
 unemployment_data
```
* Joining datasets
```{r}
#Join pols_month_data and snp_data together by a concatenated primary key year and month
fivethirtyeight_first_data = 
  left_join(pols_month_data, snp_data, by = c("year", "month")) %>%
#Ensure the names stay clean
  janitor::clean_names()

#Join the dataset above with unemployment data by a concatenated primary key year and month and make the resulting dataset look nice
fivethirtyeight_data = 
  left_join(fivethirtyeight_first_data, unemployment_data, by = c("year", "month")) %>%
  janitor::clean_names()

#The resulting dataset has all three datasets from problem 2 by year and month
fivethirtyeight_data
```
These data show various election results from months and years along with certain SNP500 close values and unemployment rates. The pols_month_data data show whether or not the governor, senators, representatives, and president were democratic or republican in that month and year from April 1947 to March 1988. The pols_month_data has `r nrow(pols_month_data)` rows and `r ncol(pols_month_data)` columns. The snp_data shows the month and year and the SNP close value of how the S&P500 stock market index did from April 1950 to May 2015. The average closing point of the S&P over the 65 years is `r mean(pull(snp_data, close))`. The snp_data has `r nrow(snp_data)` rows and `r ncol(snp_data)` columns.The unemployment_data dataset shows the unemployment rate for that month and year from January 1948 to December 2015. The average unemployment rate is `r mean(pull(unemployment_data, rate))`. The unemployment_data has `r nrow(unemployment_data)` rows and `r ncol(unemployment_data)`columns. It's important to note that all three of these datasets have different starting and ending months and years.The final combined dataset, fivethirtyeight_data has `r nrow(fivethirtyeight_data)` rows and `r ncol(fivethirtyeight_data)` columns. 

# Problem 3
```{r}
#Import in popular baby names dataset
popular_baby = 
  read_csv(file = "./data_hw2/Popular_Baby_Names.csv") %>%
#Clean names
  janitor::clean_names() %>%
#Mutate the variable ethnicity to be in consistent format (changed the names to be all the same)
  mutate(ethnicity = recode(ethnicity, 'ASIAN AND PACIFIC ISLANDER' = "asian and pacific islander",'BLACK NON HISPANIC' = "black", 'HISPANIC' = "hispanic", 'WHITE NON HISPANIC' = "white", 'ASIAN AND PACI' = "asian and pacific islander", 'BLACK NON HISP' = "black", 'WHITE NON HISP' = "white")) %>%
#Mutate the childs_first_name values to lowercase as some are upper and some are lower
  mutate(childs_first_name = str_to_lower(childs_first_name)) %>%
  mutate(gender = str_to_lower(gender)) %>%
#Group the data by first name, rank, and gender
  group_by(childs_first_name, rank, gender) %>%
#Ensure the rows are distinct
  distinct() 
  
popular_baby
```


*Create a table of the popularity of the name Olivia
```{r}
#Create data table with name Olivia as female baby name over time
female_name = 
  filter(popular_baby, childs_first_name == "olivia") %>%
  select(-gender, -count, -childs_first_name) %>%
  pivot_wider(names_from = "year_of_birth",
              values_from = "rank") %>%
  knitr::kable()

female_name
```


* Create a table of the most popular male baby names
```{r}
male_name = 
  filter(popular_baby, gender == "male", rank == 1) %>%
  select(-gender, -count, -rank) %>%
  pivot_wider(names_from = "year_of_birth",
              values_from = "childs_first_name") %>%
  knitr::kable()
male_name
```


* Create a table of male, white, non-Hispanic children in 2016
```{r}
# selected the popular baby dataset and filtered for white ethnicity, male gender, and year of birth 2016
# All of these babies are non-Hispanic per the definition of the dataset and how I cleaned it earlier
white_male = 
  select(popular_baby, year_of_birth, gender, ethnicity, childs_first_name, rank, count) %>%
  filter(ethnicity == "white", gender == "male", year_of_birth == 2016)
white_male
```


* Creating ggplot for white male
```{r}
#Using this dataset white_male we plotted x as the rank and y as the count for 2016 white, non-Hispanic babies of count against rank
ggplot(white_male, aes(x = rank, y = count, color = count)) +
  geom_point()

```

